Pleiotropy of expression quantitative trait loci increases with the number of regulated genes, active cell types and tissues and bound transcription factors

### 2024-11-10 ms/ms06_backmatter.tex

    \begin{figure}[!ht]
        %
        \caption{Histogram showing the percentage of GWAS with the percentage of their tophits colocalized with eQTLs.
        I have excluded tophits in in the MHC locus in chromosome 6 between 25,000,000 and 35,000,000,
            because this locus is not included in the colocalization pipeline.}
        \label{fig:1}
        %
    \end{figure}

@@

    \begin{figure}[!ht]
        %
        \caption{Trait clustering based on the Spearman correlation between beta values of colocalized eQTL variants, genes and biosamples.}
        \label{fig:2}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 3: Histograms of variants vs GWAS, eQTL genes and eQTL biosamples
% Fig 4: Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
% Fig 5: VEP consequences
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
    \begin{figure}[!ht]
        %
        \caption{Distribution of GWAS category counts, correlation and ENSEMBL variant effect prediction.
        \textbf{a-c} Distribution of GWAS category, eQTL gene and eQTL biosample counts.
        Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eQTL genes (\textbf{b}) and eQTL tissues (\textbf{c}).
        \textbf{d} Spearman correlation between the counts of GWAS categories, eQTL genes and eQTL biosamples.
        \textbf{e} Odds ratio of the eQTLs annotated with the given ENSEMBL variant predictor
        for eQTLs association withe the given GWAS category count versus the eQTLs with a single GWAS category count.
        The significance is given for the q-value with False discovery rate 5\%.
        }
        \label{fig:3}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 6: beta and pval
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[!ht]
        %
        \caption{eQTL and GWAS beta and p-values and allele frequencies for different GWAS category counts.
        \textbf{a,b} Average eQTL and GWAS beta values.
        \textbf{c,d} Average eQTL and GWAS p-values.
        \textbf{e} Average alternative allele frequency in the European population from the 1000 Genomes database.
        }
        \label{fig:4}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 7: Plots of the number of genes and biosample per variant-biosample and variant-gene, respectively
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[!ht]
        %
        \caption{Mean number of eQTL genes per eQTL-biosample pair (\textbf{a}) and mean number of biosamples per eQTL-gene pairs (\textbf{b}).} \label{fig:gwas_egene_etisue_per_variant}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 8: TF count per GWAS class count
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[!ht]
        %
        \caption{Analysis of transcription factor binding.
        \textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic eQTLs with a radius of 50 bp as a function of the GWAS class count.
        \textbf{b}, Odds ratio of eQTLs annotated as belonging to a cis regulatory modules vs. non-annotated.}
        \label{fig:freq_tf_per_variant}
        %
    \end{figure}

>>>>

\begin{figure}[!ht]
    %
    \caption{Distribution of GWAS based on eQTL colocalization and UCSC screenshots with colocalized GWAS/eQTLs.
    \textbf{a} Histogram showing the percentage of GWAS with the percentage of their tophits colocalized with eQTLs.
    I have excluded tophits in in the MHC locus in chromosome 6 between 25,000,000 and 35,000,000,
    because this locus is not included in the colocalization pipeline.
    \textbf{b-d} UCSC screenshots of eQTLs colocalized with GWAS variants in the chr5:132,137,990-132,728,110 (hg38)
    for the b-cells (\textbf{b}), classical monocytes (\textbf{c}) and cortex cells (\textbf{d}).
    There are two tracks for each biosample: "beta equal" and "beta unequal".
    "beta equal" corresponds to colocalized GWAS/eQTLs where both the GWAS beta and the eQTL beta coefficient have the same sign.
    "beta unequal" corresponds to colocalized GWAS/eQTLs where the GWAS beta and the eQTL beta coefficients have different signs.
    This allows to infer the beta coefficient of the GWAS variant eventhough the sign is shown with the column for the eQTL.
    }
    \label{fig:1}
    %
\end{figure}

@@

\begin{figure}[!ht]
    %
    \caption{
    Trait clustering based on the Spearman correlation between beta values of colocalized eQTL variants, genes and biosamples.
    }
    \label{fig:2}
    %
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 3: Histograms of variants vs GWAS, eQTL genes and eQTL biosamples
% Fig : Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
% Fig : VEP consequences
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!ht]
    %
    \caption{Distribution of GWAS category counts, correlation and ENSEMBL variant effect prediction.
    \textbf{a-c} Distribution of GWAS category, eQTL gene and eQTL biosample counts.
    Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eQTL genes (\textbf{b}) and eQTL tissues (\textbf{c}).
    \textbf{d} Spearman correlation between the counts of GWAS categories, eQTL genes and eQTL biosamples.
    \textbf{e} Odds ratio of the eQTLs annotated with the given ENSEMBL variant predictor
    for eQTLs association withe the given GWAS category count versus the eQTLs with a single GWAS category count.
    The significance is given for the q-value with False discovery rate 5\%.
    }
    \label{fig:3}
    %
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 4: beta and pval
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!ht]
    %
    \caption{
    eQTL and GWAS beta and p-values and allele frequencies for different GWAS category counts.
    \textbf{a,b} Average eQTL and GWAS beta values.
    \textbf{c,d} Average eQTL and GWAS p-values.
    \textbf{e} Average alternative allele frequency in the European population from the 1000 Genomes database.
    }
    \label{fig:4}
    %
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 5: Plots of the number of genes and biosample per variant-biosample and variant-gene, respectively
% Fig: TF count per GWAS class count
% Fig: eQTL gene distribution
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!ht]
    %
    \caption{
    Analysis of eQTL gene and biosample counts, transcription factor binding and gene distance as a function of the GWAS category count.
    \textbf{a} Mean number of eQTL genes per eQTL-biosample pair.
    \textbf{b} Mean number of biosamples per eQTL-gene pair.
    \textbf{c} Binding count of transcription factors in the region surrounding pleiotropic eQTLs with a radius of 50 bp as a function of the GWAS class count.
    \textbf{d} Odds ratio of eQTLs annotated as belonging to a cis regulatory modules versus non-annotated.
    \textbf{e} Violin plots of the eQTL gene distances.
    }
    \label{fig:5}
    %
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Graphical conclusions
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!ht]
    %
        \caption{
        Graphical conclusions. I have analysed the differences underlying pleiotropy of eQTLs.
        I have found that pleiotropic eQTLs have more target genes and lower GWAS and expression beta coefficients.
        There is some evidence that pleitropic eQTLs overlap with more cis-regulatory modules.
        }
    \label{fig:6}
    %
\end{figure}
